### 2022-10-22 dissertation/chapters/06-bayesian-hyper-heuristic.tex
🚧 WIP - Save point (compiles)

Formal classification of the \Ac{BHH} is needed. Burke et al. \cite{ref:burke:2010} proposed a classification mechanism that was discussed in detail in Chapter \ref{chap:hhs}. The same classification mechanism is used to classify the \ac{BHH}. For the first dimension of the classification, the authors propose that the \Ac{BHH} be classified as a \textit{selection} \ac{HH}. For the second dimension of the classification proposed, the authors propose that the \Ac{BHH} be classified as utilising \textit{perturbation} low-level heuristics. Furthermore, the \ac{BHH} is also classified as a population-based approach that includes meta-heuristics. Burke et al. \cite{ref:burke:2010} further proposed the classification of \acp{HH} according to the source of the feedback used during learning. It can thus be said that the \ac{BHH} employs online learning according to this criteria, classifying it as an \textit{adaptive} \ac{HH}. A \textit{population-based, selection, meta-hyper-heuristic that utilised perturbation of low-level heuristics and online learning} seems like an mouthful, but technically, that is it's full and formal classification. A breakdown and motivation of these classifications is given below:

@@

This section aims to present the reader with all the high level components in the architecture of the \Ac{BHH}. Burke et al. \cite{ref:burke:2010} proposed an initial framework for \acp{HH} and Grobler \cite{ref:grobler:2015} further proposed a framework for a heterogeneous meta-\ac{HH}. These frameworks have been adapted for the implementation of the \Ac{BHH}. An illustration of the high-level architecture of the \Ac{BHH} is given in Figure \ref{fig:bhh_architecture} below.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.4\textwidth]{shallow_ffnn.png}
    \caption[Shallow \index{Feedforward Neural Network}Feedforward Neural Network]{An illustration of
    a shallow \index{Feedforward Neural Network}Feedforward Neural Network.}
    \label{fig:shallow_ffnn}
\end{figure}

@@

The domain barrier presents the boundary of information flow between the heuristic space and the solution space. It is mentioned in Section \ref{sec:bhh:loss_function} that the heuristic-entity loss is used by the \ac{BHH}. Notice that the \ac{BHH} uses this performance information and no information from the solution space itself. Furthermore, the \ac{BHH} maps this loss to a credit assignment problem as is mentioned by Burke et al. \cite{ref:burke:2010}. The \ac{BHH} at a high level, is not aware of the actual candidate solution that is represented by the entity, but rather just that there is a candidate solution with a unique identifier that, together with a heuristic, produced a certain level of performance. The domain barrier is a feature that occurs in al \acp{HH}.  Burke et al. \cite{ref:burke:2010} states that the \ac{HH} collects domain-independent information from the domain barrier. A number of examples of domain independent information is given, these include the number of heuristics, the changes in evaluation function, whether a new solution was found or generated, the distance between two solutions, whether it has become stuck.

@@

The credit assignment strategy is an implementation that maps a continuous performance metric into a discrete credit metric. This component implements the ``move acceptance'' process as proposed by Özcan et al. \cite{ref:ozcan:2006,ref:ozcan:2008}. It determines if a given performance is deemed good or not. This simple discrete transformation is subject to the credit assignment problem as discussed by Burke et al. \cite{ref:burke:2010}. A good credit assignment strategy will correctly allocate credit to the appropriate heuristic-entity combination. THe magnitude of this credit is to be empirically tested.  It is later shown how these credit metrics are used to accumulate ``pseudo counts'' which concentrates/biases heuristic selection towards good performing heuristics. The \ac{BHH} propose the following initial credit assignment strategies.

@@

The dependence between random events can have a drastic impact on the probabilistic model that is implemented. For simplicity, the \ac{BHH} assumes independence between events, although the event $C$ is clearly dependent on the occurrence of $H$ and $E$. Notice that the \ac{BHH} is fundamentally a classifier, classifying which heuristic to select given an entity and performance criteria. It is later shown in Section \ref{sec:bhh:selection_mechanism:naive_bayes} that the \ac{BHH} implements a Naïve Bayes classifier. Domingos et al. \cite{ref:domingos:1997} mentions that although Bayesian classifiers’ probability estimates are only optimal under quadratic loss if the independence assumption holds, the classifier itself can be optimal under zero-one loss (misclassification rate) even when this assumption is violated by a wide margin. This means that independence can be assumed when the probabilistic model is used as a classifier and not to get real probabilities. Evaluating for proportionality and not precision allows for the assumption of independence.

@@

\index{Population Size}\textit{Population size} refers to the number of \index{entity}entities included in the \index{Entity Pool}entity pool/\index{Population}population. The exact \index{Population Size}population size to use is to be determined empirically. Naturely, an assumption is made that a larger population could lead to better results, assuming that entities are initialised to be spread uniformly across the solution search space. Oldewage \cite{ref:oldewage:2017} mentions that a large number of particles (entities) may be able to traverse a greater portion of the search space as every particle provides additional information about the search space. However, it is to be determined if this is indeed the case for the \ac{BHH}. The authors propose the following considerations when picking a \index{population size}population size:

>>>>

Formal classification of the \Ac{BHH} is needed. Burke et al.~\cite{ref:burke:2010} proposed a classification mechanism that was discussed in detail in Chapter \ref{chap:hhs}. The same classification mechanism is used to classify the \ac{BHH}. For the first dimension of the classification, the authors propose that the \Ac{BHH} be classified as a \textit{selection} \ac{HH}. For the second dimension of the classification proposed, the authors propose that the \Ac{BHH} be classified as utilising \textit{perturbation} low-level heuristics. Furthermore, the \ac{BHH} is also classified as a population-based approach that includes meta-heuristics. Burke et al.~\cite{ref:burke:2010} further proposed the classification of \acp{HH} according to the source of the feedback used during learning. It can thus be said that the \ac{BHH} employs online learning according to this criteria, classifying it as an \textit{adaptive} \ac{HH}. A \textit{population-based, selection, meta-hyper-heuristic that utilised perturbation of low-level heuristics and online learning} seems like an mouthful, but technically, that is it's full and formal classification. A breakdown and motivation of these classifications is given below:

@@

This section aims to present the reader with all the high level components in the architecture of the \Ac{BHH}. Burke et al.~\cite{ref:burke:2010} proposed an initial framework for \acp{HH} and Grobler~\cite{ref:grobler:2015} further proposed a framework for a heterogeneous meta-\ac{HH}. These frameworks have been adapted for the implementation of the \Ac{BHH}. An illustration of the high-level architecture of the \Ac{BHH} is given in Figure \ref{fig:bhh_architecture} below.

@@

\begin{figure}[H]
      \centering
      \includegraphics[width=0.4\textwidth]{shallow_ffnn.pdf}
      \caption[Shallow \index{Feedforward Neural Network}Feedforward Neural Network]{An illustration of
            a shallow \index{Feedforward Neural Network}Feedforward Neural Network.}
      \label{fig:shallow_ffnn}
\end{figure}

@@

The domain barrier presents the boundary of information flow between the heuristic space and the solution space. It is mentioned in Section \ref{sec:bhh:loss_function} that the heuristic-entity loss is used by the \ac{BHH}. Notice that the \ac{BHH} uses this performance information and no information from the solution space itself. Furthermore, the \ac{BHH} maps this loss to a credit assignment problem as is mentioned by Burke et al.~\cite{ref:burke:2010}. The \ac{BHH} at a high level, is not aware of the actual candidate solution that is represented by the entity, but rather just that there is a candidate solution with a unique identifier that, together with a heuristic, produced a certain level of performance. The domain barrier is a feature that occurs in al \acp{HH}.  Burke et al.~\cite{ref:burke:2010} states that the \ac{HH} collects domain-independent information from the domain barrier. A number of examples of domain independent information is given, these include the number of heuristics, the changes in evaluation function, whether a new solution was found or generated, the distance between two solutions, whether it has become stuck.

@@

The credit assignment strategy is an implementation that maps a continuous performance metric into a discrete credit metric. This component implements the ``move acceptance'' process as proposed by Özcan et al.~\cite{ref:ozcan:2006,ref:ozcan:2008}. It determines if a given performance is deemed good or not. This simple discrete transformation is subject to the credit assignment problem as discussed by Burke et al.~\cite{ref:burke:2010}. A good credit assignment strategy will correctly allocate credit to the appropriate heuristic-entity combination. THe magnitude of this credit is to be empirically tested.  It is later shown how these credit metrics are used to accumulate ``pseudo counts'' which concentrates/biases heuristic selection towards good performing heuristics. The \ac{BHH} propose the following initial credit assignment strategies.

@@

The dependence between random events can have a drastic impact on the probabilistic model that is implemented. For simplicity, the \ac{BHH} assumes independence between events, although the event $C$ is clearly dependent on the occurrence of $H$ and $E$. Notice that the \ac{BHH} is fundamentally a classifier, classifying which heuristic to select given an entity and performance criteria. It is later shown in Section \ref{sec:bhh:selection_mechanism:naive_bayes} that the \ac{BHH} implements a Naïve Bayes classifier. Domingos et al.~\cite{ref:domingos:1997} mentions that although Bayesian classifiers’ probability estimates are only optimal under quadratic loss if the independence assumption holds, the classifier itself can be optimal under zero-one loss (misclassification rate) even when this assumption is violated by a wide margin. This means that independence can be assumed when the probabilistic model is used as a classifier and not to get real probabilities. Evaluating for proportionality and not precision allows for the assumption of independence.

@@

\index{Population Size}\textit{Population size} refers to the number of \index{entity}entities included in the \index{Entity Pool}entity pool/\index{Population}population. The exact \index{Population Size}population size to use is to be determined empirically. Naturely, an assumption is made that a larger population could lead to better results, assuming that entities are initialised to be spread uniformly across the solution search space. Oldewage~\cite{ref:oldewage:2017} mentions that a large number of particles (entities) may be able to traverse a greater portion of the search space as every particle provides additional information about the search space. However, it is to be determined if this is indeed the case for the \ac{BHH}. The authors propose the following considerations when picking a \index{population size}population size:

### 2022-10-22 dissertation/chapters/07-methodology.tex
🚧 WIP - Save point (compiles)

This section provides the detail around the different datasets that are used throughout the empirical process. These datasets originate from the UCI Machine Learning Repository \cite{ref:uci:2022}. Datasets are grouped by problem type and include 8 classification and 7 regression datasets. The classification datasets are given in Table \ref{tab:methodology:datasets:classification} and the regression datasets are given in Table \ref{tab:methodology:datasets:regression} below.

\begin{table}[htbp]
    \centering
    \caption{Classification datasets}
    \label{tab:methodology:datasets:classification}%
	\par\bigskip
	\resizebox{\textwidth}{!}{
        \begin{tabular}{ccccccccc}
            \textbf{dataset} & \textbf{output} & \textbf{types} & \textbf{attributes} & \textbf{classes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation} \\
            \midrule
            iris  & multivariate & real  & 4     & 3     & 150   & 16    & 10    & \cite{ref:fisher:1936} \\
            car   & multivariate & categorical & 6     & 4     & 1728  & 128   & 14    & \cite{ref:bohanec:1988} \\
            abalone & multivariate & categorical, integer, real & 8     & 28    & 4177  & 256   & 17    & \cite{ref:waugh:1995} \\
            mushroom & multivariate & categorical & 22    & 2     & 8214  & 512   & 17    & \cite{ref:schlimmer:1987} \\
            wine\_quality & multivariate & real  & 12    & 11    & 4898  & 256   & 20    & \cite{ref:cortez:2009} \\
            bank  & multivariate & real  & 17    & 2     & 45211 & 512   & 89    & \cite{ref:moro:2014} \\
            diabetic & multivariate & integer & 55    & 3     & 100000 & 1024  & 98    & \cite{ref:strack:2014} \\
            adult & multivariate & categorical, integer & 14    & 2     & 48842 & 256   & 191   & \cite{ref:kohavi:1996} \\
        \end{tabular}%
    }
\end{table}%

\begin{table}[htbp]
    \centering
    \caption{Regression datasets}
    \label{tab:methodology:datasets:regression}%
	\par\bigskip
	\resizebox{\textwidth}{!}{
        \begin{tabular}{cccccccc}
            \textbf{dataset} & \textbf{output} & \textbf{types} & \textbf{attributes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation} \\
            \midrule
            fish\_toxicity & multivariate & real  & 7     & 908   & 64    & 15    & \cite{ref:cassotti:2015} \\
            housing & univariate & real  & 13    & 506   & 32    & 16    & \cite{ref:harrison:1978} \\
            forest\_fires & multivariate & real  & 13    & 517   & 32    & 17    & \cite{ref:cortez:2007} \\
            student\_performance & multivariate & integer & 33    & 649   & 32    & 21    & \cite{ref:cortez:2008} \\
            parkinsons & multivariate & integer, real & 26    & 5875  & 256   & 23    & \cite{ref:tsanas:2009} \\
            air\_quality & multivariate, timeseries & real  & 15    & 9358  & 256   & 37    & \cite{ref:de:2008} \\
            bike  & univariate & integer, real & 16    & 17389 & 256   & 68    & \cite{ref:fanaee:2014} \\
        \end{tabular}%
    }
\end{table}%

@@

        \begin{tabular}{llll}
            \textbf{heuristic} & \textbf{configuration} & \textbf{value} & \textbf{citation} \\
            \midrule
            sgd   & learning\_rate & 0.1**0.01 & \cite{ref:sutskever:2013} \\
            momentum & learning\_rate & 0.1**0.01 & \cite{ref:sutskever:2013} \\
                  & momentum & 0.9   &  \\
            nag   & learning\_rate & 0.1**0.01 & \cite{ref:sutskever:2013} \\
                  & momentum & 0.9   &  \\
            adagrad & learning\_rate & 0.1**0.01 & \cite{ref:duchi:2011} \\
                  & epsilon & 1E-07 &  \\
            rmsprop & learning\_rate & 0.1**0.01 & \cite{ref:hinton:2012} \\
                  & rho   & 0.95  &  \\
                  & epsilon & 1E-07 &  \\
            adadelta & learning\_rate & 1.0**0.95 & \cite{ref:zeiler:2012} \\
                  & rho   & 0.95  &  \\
                  & epsilon & 0.0000001 &  \\
            adam  & learning\_rate & 0.1**0.01 & \cite{ref:kingma:2014} \\
                  & beta1 & 0.9   &  \\
                  & beta2 & 0.95  &  \\
                  & epsilon & 1E-07 &  \\
            pso   & population\_size & 10    & \cite{ref:van:2010} \\
                  & learning\_rate & 0.1**0.01 &  \\
                  & inertia\_weight (w) & 0.729844 &  \\
                  & cognitive\_control (c1) & 1.49618 &  \\
                  & social\_control (c2) & 1.49618 &  \\
                  & velocity\_clip\_min & -1.0  &  \\
                  & velocity\_clip\_max & 1.0   &  \\
            de    & population\_size & 10    & \cite{ref:mezura:2006} \\
                  & selection\_strategy & best  &  \\
                  & xo\_strategy & exp   &  \\
                  & recombination probability & 0.9**0.1 &  \\
                  & beta  & 2.0**0.1 &  \\
            ga    & population\_size & 10    & \cite{ref:lambora:2019} \\
                  & selection\_strategy & rand  &  \\
                  & xo\_strategy & bin   &  \\
                  & mutation\_rate & 0.2**0.05 &  \\
        \end{tabular}%
    }
\end{table}%

@@

Two different configurations of the \ac{BHH} baseline is trained using the iris dataset \cite{ref:fisher:1936}. These configurations include a replay window size of 10 (short memory) and a high replay window size of 250 (long memory). Each configuration is repeat 3 times. Finally these configurations are compared to a run that is purely random.

>>>>

This section provides the detail around the different datasets that are used throughout the empirical process. These datasets originate from the UCI Machine Learning Repository~\cite{ref:uci:2022}. Datasets are grouped by problem type and include 8 classification and 7 regression datasets. The classification datasets are given in Table \ref{tab:methodology:datasets:classification} and the regression datasets are given in Table \ref{tab:methodology:datasets:regression} below.

\begin{table}[htbp]
      \centering
      \caption{Classification datasets}
      \label{tab:methodology:datasets:classification}%
      \par\bigskip
      \resizebox{\textwidth}{!}{
            \begin{tabular}{ccccccccc}
                  \textbf{dataset} & \textbf{output} & \textbf{types}             & \textbf{attributes} & \textbf{classes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}          \\
                  \midrule
                  iris             & multivariate    & real                       & 4                   & 3                & 150                & 16             & 10             & ~\cite{ref:fisher:1936}    \\
                  car              & multivariate    & categorical                & 6                   & 4                & 1728               & 128            & 14             & ~\cite{ref:bohanec:1988}   \\
                  abalone          & multivariate    & categorical, integer, real & 8                   & 28               & 4177               & 256            & 17             & ~\cite{ref:waugh:1995}     \\
                  mushroom         & multivariate    & categorical                & 22                  & 2                & 8214               & 512            & 17             & ~\cite{ref:schlimmer:1987} \\
                  wine\_quality    & multivariate    & real                       & 12                  & 11               & 4898               & 256            & 20             & ~\cite{ref:cortez:2009}    \\
                  bank             & multivariate    & real                       & 17                  & 2                & 45211              & 512            & 89             & ~\cite{ref:moro:2014}      \\
                  diabetic         & multivariate    & integer                    & 55                  & 3                & 100000             & 1024           & 98             & ~\cite{ref:strack:2014}    \\
                  adult            & multivariate    & categorical, integer       & 14                  & 2                & 48842              & 256            & 191            & ~\cite{ref:kohavi:1996}    \\
            \end{tabular}%
      }
\end{table}%

\begin{table}[htbp]
      \centering
      \caption{Regression datasets}
      \label{tab:methodology:datasets:regression}%
      \par\bigskip
      \resizebox{\textwidth}{!}{
            \begin{tabular}{cccccccc}
                  \textbf{dataset}     & \textbf{output}          & \textbf{types} & \textbf{attributes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}         \\
                  \midrule
                  fish\_toxicity       & multivariate             & real           & 7                   & 908                & 64             & 15             & ~\cite{ref:cassotti:2015} \\
                  housing              & univariate               & real           & 13                  & 506                & 32             & 16             & ~\cite{ref:harrison:1978} \\
                  forest\_fires        & multivariate             & real           & 13                  & 517                & 32             & 17             & ~\cite{ref:cortez:2007}   \\
                  student\_performance & multivariate             & integer        & 33                  & 649                & 32             & 21             & ~\cite{ref:cortez:2008}   \\
                  parkinsons           & multivariate             & integer, real  & 26                  & 5875               & 256            & 23             & ~\cite{ref:tsanas:2009}   \\
                  air\_quality         & multivariate, timeseries & real           & 15                  & 9358               & 256            & 37             & ~\cite{ref:de:2008}       \\
                  bike                 & univariate               & integer, real  & 16                  & 17389              & 256            & 68             & ~\cite{ref:fanaee:2014}   \\
            \end{tabular}%
      }
\end{table}%

@@

            \begin{tabular}{llll}
                  \textbf{heuristic} & \textbf{configuration}    & \textbf{value} & \textbf{citation}          \\
                  \midrule
                  sgd                & learning\_rate            & 0.1**0.01      & ~\cite{ref:sutskever:2013} \\
                  momentum           & learning\_rate            & 0.1**0.01      & ~\cite{ref:sutskever:2013} \\
                                     & momentum                  & 0.9            &                            \\
                  nag                & learning\_rate            & 0.1**0.01      & ~\cite{ref:sutskever:2013} \\
                                     & momentum                  & 0.9            &                            \\
                  adagrad            & learning\_rate            & 0.1**0.01      & ~\cite{ref:duchi:2011}     \\
                                     & epsilon                   & 1E-07          &                            \\
                  rmsprop            & learning\_rate            & 0.1**0.01      & ~\cite{ref:hinton:2012}    \\
                                     & rho                       & 0.95           &                            \\
                                     & epsilon                   & 1E-07          &                            \\
                  adadelta           & learning\_rate            & 1.0**0.95      & ~\cite{ref:zeiler:2012}    \\
                                     & rho                       & 0.95           &                            \\
                                     & epsilon                   & 0.0000001      &                            \\
                  adam               & learning\_rate            & 0.1**0.01      & ~\cite{ref:kingma:2014}    \\
                                     & beta1                     & 0.9            &                            \\
                                     & beta2                     & 0.95           &                            \\
                                     & epsilon                   & 1E-07          &                            \\
                  pso                & population\_size          & 10             & ~\cite{ref:van:2010}       \\
                                     & learning\_rate            & 0.1**0.01      &                            \\
                                     & inertia\_weight (w)       & 0.729844       &                            \\
                                     & cognitive\_control (c1)   & 1.49618        &                            \\
                                     & social\_control (c2)      & 1.49618        &                            \\
                                     & velocity\_clip\_min       & -1.0           &                            \\
                                     & velocity\_clip\_max       & 1.0            &                            \\
                  de                 & population\_size          & 10             & ~\cite{ref:mezura:2006}    \\
                                     & selection\_strategy       & best           &                            \\
                                     & xo\_strategy              & exp            &                            \\
                                     & recombination probability & 0.9**0.1       &                            \\
                                     & beta                      & 2.0**0.1       &                            \\
                  ga                 & population\_size          & 10             & ~\cite{ref:lambora:2019}   \\
                                     & selection\_strategy       & rand           &                            \\
                                     & xo\_strategy              & bin            &                            \\
                                     & mutation\_rate            & 0.2**0.05      &                            \\
            \end{tabular}%
      }
\end{table}%

@@

Two different configurations of the \ac{BHH} baseline is trained using the iris dataset~\cite{ref:fisher:1936}. These configurations include a replay window size of 10 (short memory) and a high replay window size of 250 (long memory). Each configuration is repeat 3 times. Finally these configurations are compared to a run that is purely random.
